\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-20 notes/2loopNLSM.tex
Lorenz gauge concern

We can achieve this construction by splitting the gauge group of the full Yang-Mills action from $G$ to $G_L \times G_R$, where the full non-abelian gauge symmetry is restored only when we identify $G_L = G_R$. However, if we break the left gauge group down to $G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to $G_R \rightarrow U(N)$, then we are left with a semi-abelian theory that is gauge invariant in the preferred $U(1)$-directions. We call this manifestly color-dual theory, \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c .
\end{equation}
Here we can think of the abelian vector as a background field that sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix \cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the heart of YZ-theory \cite{Cheung:2016prv}, J-theory \cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills \cite{Monteiro2011pc}.
Specifically, semi-abelian YM is simply a clever reinterpretation of the Lagrangian obtained by integrating an auxiliary field into the J-theory equations of motion.
The Feynman rule for the cubic vertex is
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
The propagator kinetic mixes between the two gauge fields and has the standard Feynman rule of $i/k^2$.
We can use the Ward identity of \cref{eq:GIrelA} to check that the amplitudes are indeed gauge invariant when the abelian polarizations are taken to be longitudinal, $\bar{\varepsilon} \rightarrow k$.

We can select out both of these manifestly cubic sectors by selecting only $\varepsilon^+\varepsilon^+$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and $YZ$-theory select with their on-shell states. Indeed, the $D$-dimensional description of $J$-theory is just the following blocks in the expansion of Yang-Mills given in \cite{Pavao:2022kog}
\begin{equation}
A(...,\bar{A}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\end{equation}
and similarly so at one-loop
\begin{equation}
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}\, .
\end{equation}
To recover NLSM at one-loop, we need to extract the $D$-dependent part that corresponds to an internal $YY$-loop from extra-dimensional scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all, $J$-theory and $YZ$-theory both have well defined propagators for internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we noted in the previous section, all the $J$ and $Z$ states must be on-shell in order to produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at one-loop will not produce NLSM amplitudes. However, the internal $YY$-loop is a valid forward limit, since as constructed YZ theory is valid for off-shell Y-particles, and thus they serve as a valid forward limit. We will construct these one-loop integrands below in the next section.

>>>>

We can achieve this construction by splitting the gauge group of the full Yang-Mills action from $G$ to $G_L \times G_R$, where the full non-abelian gauge symmetry is restored only when we identify $G_L = G_R$. However, if we break the left gauge group down to $G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to $G_R \rightarrow U(N)$, then we are left with a semi-abelian theory that is gauge invariant in the preferred $U(1)$-directions. We call this manifestly color-dual theory, \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
where $A_\mu$ is in Lorenz gauge.
\jm{I'm pretty sure A has to be in Lorenz gauge.}
Here we can think of the abelian vector as a background field that sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix \cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the heart of YZ-theory \cite{Cheung:2016prv}, J-theory \cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills \cite{Monteiro2011pc}.
Specifically, semi-abelian YM is simply a clever reinterpretation of the Lagrangian obtained by integrating an auxiliary field into the J-theory equations of motion.
The Feynman rule for the cubic vertex is
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
The propagator kinetic mixes between the two gauge fields and has the standard Feynman rule of $i/k^2$.
We can use the Ward identity of \cref{eq:GIrelA} to check that the amplitudes are indeed gauge invariant when the abelian polarizations are taken to be longitudinal, $\bar{\varepsilon} \rightarrow k$.

\jm{Do you need to put semi-Abelian YM in axial/lightcone gauge in order to recover SDYM?}
We can select out both of these manifestly cubic sectors by selecting only $\varepsilon^+\varepsilon^+$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and $YZ$-theory select with their on-shell states. Indeed, the $D$-dimensional description of $J$-theory is just the following blocks in the expansion of Yang-Mills given in \cite{Pavao:2022kog}
\begin{equation}
A(...,\bar{A}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\end{equation}
and similarly so at one-loop
\begin{equation}
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}\, .
\end{equation}
To recover NLSM at one-loop, we need to extract the $D$-dependent part that corresponds to an internal $YY$-loop from extra-dimensional scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all, $J$-theory and $YZ$-theory both have well defined propagators for internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we noted in the previous section, all the $J$ and $Z$ states must be on-shell in order to produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at one-loop will not produce NLSM amplitudes. However, the internal $YY$-loop is a valid forward limit, since as constructed YZ theory is valid for off-shell Y-particles, and thus they serve as a valid forward limit. We will construct these one-loop integrands below in the next section.
